# Visualizing data {#visualdata}

### Plotting candlestick chart with technicals 

The function **create_candlestick()** takes 3 arguments: **a ticker**, a **start date** (by default, it is a year ago), an **end date** (by default, it is today).  We got the main part of the function from the **tidyquant** [package](https://github.com/business-science/tidyquant).  
```{r plot_candlestick}
geom_candlestick <- function(mapping = NULL, data = NULL, stat = "identity",
                                position = "identity", na.rm = TRUE, show.legend = NA,
                                inherit.aes = TRUE,
                                colour_up = "gray30", colour_down = "gray30",
                                fill_up = "green3", fill_down = "red",
                                ...) {

    linerange <- ggplot2::layer(
        stat = StatLinerangeBC, geom = GeomLinerangeBC, data = data, mapping = mapping,
        position = position, show.legend = show.legend, inherit.aes = inherit.aes,
        params = list(na.rm = na.rm, fill_up = fill_up, fill_down = fill_down,
                      colour_up = colour_up, colour_down = colour_down, ...)
    )

    rect <- ggplot2::layer(
        stat = StatRectCS, geom = GeomRectCS, data = data, mapping = mapping,
        position = position, show.legend = show.legend, inherit.aes = inherit.aes,
        params = list(na.rm = na.rm, fill_up = fill_up, fill_down = fill_down,
                      colour_up = colour_up, colour_down = colour_down, ...)
    )

    list(linerange, rect)
}

StatLinerangeBC <- ggplot2::ggproto("StatLinerangeBC", Stat,
                                    required_aes = c("x", "open", "high", "low", "close"),

                                    compute_group = function(data, scales, params,
                                                             fill_up, fill_down,
                                                             colour_up, colour_down) {

                                        data <-  data %>%
                                            dplyr::mutate(colour = ifelse(open < close, colour_up, colour_down))

                                        tibble::tibble(x = data$x,
                                                       ymin = data$low,
                                                       ymax = data$high,
                                                       colour = data$colour)
                                    })


StatRectCS <- ggplot2::ggproto("StatRectCS", Stat,
                                required_aes = c("x", "open", "high", "low", "close"),

                                compute_group = function(data, scales, params,
                                                         fill_up, fill_down,
                                                         colour_up, colour_down) {

                                    data <-  data %>%
                                        dplyr::mutate(fill = ifelse(open < close, fill_up, fill_down),
                                                      ymin = ifelse(open < close, open, close),
                                                      ymax = ifelse(open < close, close, open))

                                    tibble::tibble(xmin = data$x - 0.45,
                                                   xmax = data$x + 0.45,
                                                   ymin = data$ymin,
                                                   ymax = data$ymax,
                                                   fill = data$fill)
                                })


GeomRectCS <- ggproto("GeomRectCS", GeomRect,
                      default_aes = aes(colour = NA,
                                        size = 0.5,
                                        linetype = 1,
                                        alpha = NA))


GeomLinerangeBC <- ggproto("GeomLinerangeBC", GeomLinerange,
                           default_aes = aes(size = 0.5,
                                             linetype = 1,
                                             alpha = NA))

create_candlestick <- function(tickerss, start_d = today()-360, end_d = today()) {
  df <- read_csv(paste0(thePath, "/stockdata_av/", tickerss, ".csv")) %>% 
    select(Index, Open, High, Low, Close, Volume)
  
  df$Index <- ymd(df$Index)
  
  #for the ADX
  yo <- TTR::ADX(df[,3:5], n = 13) %>% as_tibble() %>% select(-DX)
  
  # Adding the other variables such a moving averages and relative strength index
  df3 <- df %>% 
    mutate(sma200 = TTR:: SMA(Close, 200), 
           sma50 = TTR::SMA(Close, 50), 
           ema9 = TTR::EMA(Close, 9), 
           rsi14 = TTR::RSI(Close, 14), 
           rsi5 = TTR::RSI(Close, 5), 
           ppo_line = (TTR::EMA(Close, n = 12) - TTR::EMA(Close, n = 26)) / TTR::EMA(Close, n = 26) * 100, 
           ppo_signal = TTR::EMA(ppo_line, n = 9)) 
  
  df2 <- bind_cols(df3, yo) %>% 
    filter(Index >= start_d & Index <= end_d)
  
  
  # The main chart with the moving averages
  p1 <- ggplot(df2, aes(x=Index, y = Close)) + 
    geom_candlestick(aes(open = Open, high = High, low = Low, close = Close)) + 
    geom_line(aes(y = ema9), color = "red", size = 0.2) + 
    geom_line(aes(y = sma200), color = "darkorchid1", size = 0.3) + 
    # because I need to remember which chart is it (to which stock it belongs)
    annotate("text", x = df2$Index[10], y = 1.1 * df2$Close[10], label = tickerss, color = "white") + 
    geom_line(aes(y = sma50), color = "Turquoise 1", size = 0.3) + 
    scale_x_bd(business.dates=df2$Index, max.major.breaks = 20, labels=date_format("%b '%y"), 
               expand = c(0,0.3)) + 
    scale_y_continuous(sec.axis = sec_axis(~.*1)) + 
    theme(axis.title.x = element_blank(), 
          axis.text.x = element_blank(), 
          axis.text.y = element_text(angle = 90), 
          # This is to shrink the padding at the 4 side of the graph
          plot.margin = margin(0.2, 0.2, 0.1, 0.4, "cm"),     
          panel.background = element_rect(fill = "black"), 
          plot.background = element_rect(fill = "Gray 65"), 
          panel.grid.major.x = element_line(color = "white", linetype = "dotted", size = 0.2), 
          panel.grid.major.y = element_line(color = "white", linetype = "dotted", size = 0.2),
          panel.grid.minor.y = element_line(color = "white", linetype = "dotted", size = 0.15),
          panel.grid.minor.x = element_blank())
  
  # graphing of the ppo part.  
  p2 <- ggplot(df2, aes(x = Index)) + 
    geom_line(aes(y = ppo_signal, color = "darkorchid1"), size = 0.4) + 
    geom_line(aes(y = ppo_line, color = "Royal Blue 1"), size = 0.5) + 
    geom_hline(yintercept = 0, color = "red", linetype = "dashed", size = 0.3) + 
    scale_y_continuous(sec.axis = sec_axis(~.*1)) + 
    scale_x_bd(business.dates=df2$Index, max.major.breaks = 20, labels=date_format("%b '%y"), 
               expand = c(0,0.1)) + 
    ylab("PPO") + 
    theme(legend.position = "none",  
          axis.title.x = element_blank(), 
          axis.text.x = element_blank(), 
          axis.text.y = element_text(angle = 90), 
          plot.margin = margin(0, 0.2, 0.1, 0.4, "cm"), 
          panel.background = element_rect(fill = "black"), 
          plot.background = element_rect(fill = "Gray 65"), 
          panel.grid.major.x = element_line(color = "white", linetype = "dotted", size = 0.2), 
          panel.grid.major.y = element_line(color = "white", linetype = "dotted", size = 0.1),
          panel.grid.minor = element_blank())
  
  p4 <- ggplot(df2, aes(x = Index)) + 
    geom_line(aes(y = DIp), color = "Turquoise 1", size = 0.2) + 
    geom_line(aes(y = DIn), color = "red", size = 0.2) + 
    geom_line(aes(y = ADX), color = "Gray 70", size = 0.3) + 
    scale_y_continuous(sec.axis = sec_axis(~.*1)) + 
    scale_x_bd(business.dates=df2$Index, max.major.breaks = 20, labels=date_format("%b '%y"), 
               expand = c(0,0.1)) + 
    ylab("ADX") + 
    theme(legend.position = "none",  
          axis.title.x = element_blank(), 
          axis.text.x = element_blank(), 
          axis.text.y = element_text(angle = 90), 
          plot.margin = margin(0, 0.2, 0.1, 0.4, "cm"), 
          panel.background = element_rect(fill = "black"), 
          plot.background = element_rect(fill = "Gray 65"), 
          panel.grid.major.x = element_line(color = "white", linetype = "dashed", size = 0.1), 
          panel.grid.major.y = element_line(color = "white", linetype = "dotted", size = 0.1),
          panel.grid.minor = element_blank()) 
  
  
  p3 <- ggplot(df2, aes(x = Index)) + 
    geom_line(aes(y = rsi14, color = "Dark Orange")) + 
    geom_line(aes(y = rsi5, color = "Gray 80"), linetype = "dotted", size = 0.4) + 
    scale_x_bd(business.dates=df2$Index, max.major.breaks = 20, labels=date_format("%b '%y"), 
               expand = c(0,0.5)) + 
    scale_y_continuous(sec.axis = sec_axis(~.)) + 
    ylab("RSI") + 
    theme(legend.position = "none", 
          axis.title.x = element_blank(),
          axis.text.y = element_text(angle = 90), 
          axis.text.x = element_text(angle = 30, vjust = 0.8), 
          plot.margin = margin(0.0, 0.2, 0.2, 0.4, "cm"), 
          panel.background = element_rect(fill = "black"), 
          plot.background = element_rect(fill = "Gray 65"), 
          panel.grid.major.x = element_line(color = "white", linetype = "dashed", size = 0.2), 
          panel.grid.major.y = element_line(color = "white", linetype = "dotted", size = 0.2),
          panel.grid.minor = element_blank())
  
  
  grid.arrange(p1, p2, p4, p3, ncol = 1, heights = c(2.5, 1, 1, 1))
}
```

Let's try it.  
```{r get_graph_data, message = FALSE, warning = FALSE, fig.width = 12, fig.height=8}
ticker <- "PAAS"
#get_eod_data(ticker)
create_candlestick(ticker, "2017-06-01", "2019-01-20")
```

### Plotting cumulative return 

This graph is useful when comparing the returns of several assets together. 
```{r cumul_return_graph, message = FALSE, warning = FALSE}
ticker1 <- "SLB"
ticker2 <- "HAL"
starting_time <- "2018-01-01"

#get_eod_data(c(ticker1, ticker2))

df1 <- read_csv(paste0(thePath, "/stockdata_av/", ticker1, ".csv")) %>% 
  select(Index, Adjusted) %>% rename(SLB = Adjusted)
df2 <- read_csv(paste0(thePath, "/stockdata_av/", ticker2, ".csv")) %>% 
  select(Index, Adjusted) %>% rename(HAL = Adjusted)

df <- left_join(df1, df2, by="Index") %>% filter(Index >= starting_time) %>% 
  mutate(slb_ret = SLB / first(SLB), hal_ret  = HAL / first(HAL)) %>% na.omit() %>% 
  select(Index, slb_ret, hal_ret) %>% gather(key = "ticker", value = "return", -Index)

ggplot(df, aes(x = Index, y = return, col = ticker)) + geom_line()



```

